Formal Analysis of EDHOC Key Establishment for Constrained IoT Devices

### 2022-07-21 formalization.tex

As in~\cite{Norr21}, we verify \mEdhoc{} in the Dolev-Yao model~\cite{DY83}.
%
This is a well-established model for symbolic verification of security protocols.
%
Messages are modelled as terms in an algebra, and the various cryptographic operations are assumed to be perfect, i.e. encrypted messages can only be decrypted with the correct key, no hash collisions exist etc.
%

The adversary is assumed to be the network, and thus controls the communication channel and can see all messages being communicated as part of the protocol.
%
In addition, they can interact with an unbounded number of protocol sessions, dropping, injecting, and modifying messages at will, but cannot break cryptography.

On top of the Dolev-Yao model, we allow the adversary two extra capabilities with respect to getting access to the long-term and ephemeral keys. 
%
Long-term key reveal, denoted $\mRevLTK^{t}(A)$, stands for the adversary
gaining access to a party $A$'s long-term private key \mPriv{A} at a time point $t$.
%
Ephemeral key reveal, denoted $\mRevEph^{t}(A, k)$, stands for the adversary obtaining, at time $t$, the ephemeral private key \mPrivE{A} used by party $A$  in a session where they establish key material $k$.
%
Ephemeral keys might be stored in a less secure part of a device, but long-term keys are often stored and operate within a secure model, like a trusted execution environment (TEE). \vnote{Do we want to say more about TEEs here?}
%
Formalizing these two events allows us to have more fine-grained control over the access that an adversary has to these fundamentally different kinds of keys.
%

We now briefly describe \mTamarin{}, and how we model \mEdhoc{} in it here.

%-------------------------------------------------------------------------- sub
\subsection{\mTamarin{}}
\label{sec:tamarin}
As in~\cite{Norr21}, we model and verify \mEdhoc{} in the symbolic model using \mTamarin{}~\cite{DBLP:conf/cav/MeierSCB13}, which is an interactive tool for the formal verification of security protocols.
%
Protocols are modelled in \mTamarin{} as multi-set rewrite rules which encode a transition relation.
%
The elements of these multi-sets, called facts, contribute to the global system state.
%
For syntactic sugar, \mTamarin{} also allows the use of let-bindings and tuples. 
%
For ease of presentation, we will present the model and properties in a slightly different syntax in this paper, but this syntax can be directly mapped to that of \mTamarin.

Rewrite rules can be annotated with events, called actions in \mTamarin{}.
%
Communicated messages in the protocol are modelled as terms in an algebra, which specifies sets of names, variables, and allowable function symbols.
%
Facts and actions are modelled as $n$-ary predicates in the term algebra, and actions can be parametrized using terms.

An annotated multi-set rewrite rule is represented as $l \ifarrow[e] r$, where $l$ and $r$ are multi-sets, and $e$ is a multi-set of actions.
%
A sequence of actions yields a protocol execution trace.
%
Event types are predicates over global states generated during protocol execution.
%
Consider an event type $E$ and a timestamp $t$ as part of a trace.
%
By $E^{t}(p_i)_{i\in\mathbb{N}}$, we represent an event of type $E$ occurring at time $t$ in a trace, parametrized by the sequence of values $(p_i)_{i\in\mathbb{N}}$ (corresponding to the action fact $E(p_i)_{i\in\mathbb{N}}@t$ in \mTamarin).
%
Thus, the time points form a quasi order, and we denote the fact that $t_{1}$ comes before $t_{2}$ in a protocol trace by $t_{1} \lessdot t_{2}$, and that $t_{1}$ and $t_{2}$ stand for the same time point in a trace by $t_{1} \doteq t_{2}$.
%
\mTamarin{} allows events to occur at the same time point, with one restriction: multiple events of the same type cannot occur simultaneously, so if $t_{1} \doteq t_{2}$, then $E^{t_{1}} = E^{t_{2}}$.

Properties are defined as formulas in a fragment of temporal first order logic, and these formulas can be verified over execution traces.

@@

Users can extend the default term algebra and equational theory in \mTamarin{} with new function symbols and unification rules for these new symbols.
%
For example, \mEdhoc{} requires authenticated encryption, which we model using the symbol \mT{aeadEncrypt}. 
%
We augment \mTamarin{} with the following rule for this operation, which represents the fact that if the adversary knows a key \mT{k}, a message \mT{m}, and authenticated data \mT{ad}, and has access to an encryption algorithm \mT{ai}, then they can obtain the message corresponding to the authenticated encryption of \mT{m} with \mT{k}~\cite{Norr21}.
\begin{verbatim}
[!KU(k), !KU(m), !KU(ad), !KU(ai)] --[]-> [!KU(aeadEncrypt(k, m, ad, ai))]
\end{verbatim}
%
Similarly, we can extend \mTamarin{} with rules for other operations, like \mXor{}.

%
The rules modelling a Dolev-Yao adversary and the evolution of their knowledge as the protocol executes are built into \mTamarin.
%
We can extend and/or modify this behaviour by adding other rules as well.
%
To denote that the adversary has access to a message $p$ at time $t$, we use $\mK^{t}(p)$.
%
This stands for the fact that as the adversary interacts with the parties executing the protocol, at time point $t$, the adversary can derive the term $p$ using the Dolev-Yao deduction rules and any additional adversary capabilities specified in \mTamarin.
%
As an example, consider the following implication 
\[
    \forall t, k, k'\mLogicDot \mK^{t}(\langle k, k'\rangle)\ \rightarrow\ 
\mK^{t}(k) \land \mK^{t}(k'),
\]
which represents the fact that if the adversary gets to know the pair of keys $\langle k, k' \rangle$ at a time point $t$, then the adversary also knows each of those keys $k$ and $k'$ at time point $t$ as well.
%
For more details about how \mTamarin{} manages adversary knowledge, see~\cite{DBLP:conf/cav/MeierSCB13}.
%

%-------------------------------------------------------------------------- sub

\subsection{Model and Desired Properties}
\label{sec:desired-properties}
In this section, we describe how we model \mEdhoc{} and the various properties we want it to satisfy.
%

@@

%
In this work, we consider the following properties: perfect forward secrecy, injective agreement for both the initiator and the responder, and implicit agreement. 
%
Agreement is considered on a set of parameters $S$ which also contains the session key material \mSessKey{}.
%
We will first describe in detail both properties, and then describe the contents of the set $S$.
%
We formalize these properties as shown in Figure~\ref{fig:props}, which is a modified version of a figure in~\cite{Norr21}.
\vnote{The figure changes because of TEEs}

@@

%----------------------------------------------------------------------- PFS
\subsubsection{Perfect Forward Secrecy (PFS): }
\label{sec:secrecy}
Informally, by perfect forward secrecy, we mean that the adversary cannot gain access to the session key material \mSessKey{}, even if a long-term key is leaked after the establishment of \mSessKey{}.
%
PFS for session key material \mSessKey{} is formalized as \mPredPfs{} in Figure~\ref{fig:props}.

In the \mIComplete{} event, $I$ represents the identity of the initiator, and $R$ represents the identity of the party who the initiator believes is playing the responder role, while \mSessKey{} stands for the established session key material.
%
\mRComplete{} has analogous parameters; in particular, the responder $R$ believes $I$ is the party playing the initiator role.
%
Essentially, the property states that an adversary obtains \mSessKey{} only if a party's long-term key is compromised before their run ends, or if an ephemeral key is compromised at any time after a party's run began.
\vnote{This has to change once we add TEEs}
%
The property in Figure~\ref{fig:props} is unlike the actual \mTamarin{} lemma in one minor way: \mTamarin's syntax does not allow disjunctions to appear on the left-hand side of an implication inside a universally-quantified formula.
%
Therefore, in the \mTamarin{} code, instead of using the disjunction $\mIComplete^{t_2}(I, R, \mSessKey)\, \lor\,  \mRComplete^{t_2}(I, R, \mSessKey)$ to model the fact that either party can have completed their execution, we use a single action parametrized by the terms $I$, $R$, and \mSessKey.

%-------------------------------------------------------------------- InjAgree
\subsubsection{Authentication Properties:}
\label{sec:authenticationDef}
Following~\cite{Norr21}, we prove two different kinds of authentication properties, namely \emph{injective agreement} in the style of~\cite{DBLP:conf/csfw/Lowe97a}, and implicit agreement.
%
Injective agreement can be guaranteed to either party running the protocol.
%
For the initiator $I$, it guarantees to $I$ that whenever $I$ believes that they have completed a run with $R$ as responder, then the party $R$ has indeed executed the protocol in the role of a responder, and that this run of $I$ uniquely corresponds to one of $R$ where the set of parameters is $S$ and includes, in particular, the session key material \mSessKey{}.
%
It can be defined for $R$ in a similar manner.

%
We formalize injective agreement for the initiator role as \mPredInjI{} and for the responder role as \mPredInjR{} in Figure~\ref{fig:props}.
%
For the initiator $I$, this property represents the fact that either injective agreement (as described above) holds, or the long-term key of the party $R$ assumed to be playing the responder role has been compromised before $I$'s role finished.
%
If the adversary has managed to compromise $R$'s long-term key, they can generate any message of their liking and signed it using said key, or computed a $\mathit{MAC}_{R}$ using the long-term keys \mPubE{I}, \mPriv{R}, and an ephemeral key pair $\langle\mPrivE{R},\ \mPubE{R}\rangle$ of their choice.
%
The analogous definition holds for the responder $R$.
\vnote{Check this, and modify to deal with TEEs}

@@

The ``implicit'' in the name of the property stands for the fact that a party $A$ assumes that any party $B$ who has access to the session key material \mSessKey{} must, in fact, be the intended party, and that if $B$ is honest, $B$ will agree on a set $S$ of parameters which includes \mSessKey.
%
Implicit agreement for both roles guarantees to $A$ that $A$ is or has been involved in exactly one protocol execution with $B$, and that $B$ agrees or will agree with $A$ on $S$.
%
This property diverges from injective agreement in that upon sending the last message, $A$ concludes that if this message reaches $B$, then $A$ and $B$ agree on each other's identities and roles, and the set $S$.

Note that for implicit agreement to hold for $I$, the ephemeral keys must not be revealed, since the property relies on the fact that the intended responder is the only one who knows the session key material.
%
If the adversary has access to the ephemeral keys, they can use it along with the public keys of $I$ and $R$ to compute the session key material.
%
However, either party's long-term key can be revealed after the other party has finished their execution, since this still leaves the adversary unable to compute \mGxy{}.

Since \mTamarin{} runs out of memory to verify this property as is, we split it into two lemmas -- \mPredImpI{} for $I$ for one \mPredImpR{} for $R$.
%
Figure~\ref{fig:props} contains the definition for \mPredImpI{}.
%
\mPredImpR{} is formalized similarly, so we omit it.

%------------------------------------------------------- Agreed parameters
\subsubsection{Parameters in set $S$:}
\label{sec:agreedParams}
We now describe the set $S$ of parameters upon which the two parties obtain guarantees via the above properties.
%
The initiator $I$ gets injective and implicit agreement guarantees on the
following partial set $S_P$ of parameters~\cite{Norr21}:
\begin{itemize}
    \item the roles played by itself and its peer,
    \item responder identity,
    \item session key material (which varies depending on the \mEdhoc{} method),
    \item context identifiers \mCi{} and \mCr{}, and
    \item cipher suites \mSuites{}.
\end{itemize}
%

@@

In addition to these properties, a couple of properties can be inferred without being explicitly modelled and verified. 
%
One might be interested in checking is called Key-Compromise Impersonation (KCI)~\cite{DBLP:conf/ima/Blake-WilsonJM97}.
%
Key-Compromise Impersonation (KCI)~\cite{DBLP:conf/ima/Blake-WilsonJM97} often occurs in protocols where a party cannot be assured that the peer they are involved in a run with actually knows the session key material.
%
A KCI attack occurs when an adversary who has access to $A$'s long-term private key to make $A$ believe that they completed an execution with a peer $B$, while $B$ did not participate in said execution at all.
%
Our above notions of agreement ensure that both parties agree on each other's identity, role, and session key material.
%
Therefore, all \mEdhoc{} methods that satisfy these agreement properties also avoid KCI attacks.

Another kind of attack is an Unknown Key Share (UKS)~\cite{DBLP:conf/ima/Blake-WilsonJM97}.
%
As part of a UKS attack, a party $A$ can be made to believe that it finished an execution with a party $B$, but where the session key material is actually shared between $A$ and $C$ instead.
%
Again, due to the agreement on identities and session key material, any method that satisfies the above agreement properties also resists UKS attacks.
%
Overall, the agreement properties capture entity authentication, and satisfy any properties based on that notion.

\vnote{Is there anything else we can add thanks to the TEEs?}

%---------------------------------------------------------------------------sub
\subsection{Encoding \mEdhoc{} in \mTamarin}
\label{sec:modeling}
%
In this section, we describe how we model the \mSigSig, \mSigStat, \mStatSig, and \mStatStat{} methods of \mEdhoc{} in \mTamarin.
%
As in~\cite{Norr21}, we construct the \mTamarin{} model by utilizing the fact that all methods of \mEdhoc{} share a common underlying structure (as shown in Figure~\ref{fig:edhocFramework}).
%
We do so by writing a single specification file in the M4 macro language, which generates all the methods.
%
As in~\cite{Norr21}, we only present the \mStatSig{} method which illustrates the use of two different authentication methods.
%
The full \mTamarin{} code for all models can be found at~\cite{edhocTamarinRepo}.
%

@@

In addition to these default operations, \mEdhoc{} is built over \mHkdfExpand, \mHkdfExtract, and the \mAead{} functions.
%
We represent \mHkdfExpand by \mT{expa} and \mHkdfExtract by \mT{extr}.
%
\mAead{} operations need us to add extra equations to the underlying theory.
%
A term encrypted using \mAead{} is represented by \mT{aeadEncrypt(m, k, ad, ai)}, where \mT{m} is the underlying message, \mT{k} is the encrypting key, \mT{ad} is the additional data, and \mT{ai} is the identifier for the encryption algorithm.
%
Decryption of such a term is defined via two equations that we add to \mTamarin's theory.
%
The following equation requires the decrypting party to know the additional data \mT{ad} to decrypt this encrypted term with verification of its integrity. 
\begin{small}
\begin{verbatim}
aeadDecrypt(aeadEncrypt(m, k, ad, ai), k, ad, ai) = m.
\end{verbatim}
\end{small}
%
Only the above equation is used by honest parties, but the adversary should also be able to decrypt without having to go through the additional step of identity verification.
%
To this end, we also add the following equation, where the adversary does not need access to the additional data \mT{ad} in order to decrypt.
\begin{small}\begin{verbatim}
decrypt(aeadEncrypt(m, k, ad, ai), k, ai) = m.
\end{verbatim}\end{small}
%

@@

To model the reveal of long-term keys and ephemeral keys to an adversary, we use standard reveal rules and events of type \mRevLTK{} and \mRevEph, respectively. 
%
It is also important to keep track of the time points at which these events occur.
%
Long-term keys can be revealed on registration, even before the protocol begins.
%
However, ephemeral keys in our model can only be revealed when a party completes their role, i.e. simultaneous with events of type \mIComplete{} and \mRComplete.
%
While we did try to reveal ephemeral keys at the start of protocol execution, we could not get the verification to terminate in such a model.
\vnote{Do we still not get \mTamarin{} to terminate if we reveal ephemeral keys at the start of the protocol? With the TEEs etc?}
%
Having set out the capabilities of the adversary, we now model the execution of the honest agents' roles.

%%%%%%%%%%% START HERE
For each protocol method, we use two rules apiece for the initiator and responder -- \mT{I1}, \mT{R2}, \mT{I3}, \mT{R4}.
%
Each of these stand for one step of the protocol as executed by either party.
%
To disambiguate, we will attach the method to the rule name.
%
These four rules directly map to the event types \mIStart, \mRStart, \mIComplete, and \mRComplete, respectively.
%
We show the \mT{R2\_STAT\_SIG} rule below to illustrate the various aspects of the \mTamarin{} modelling we are describing here.

>>>>

As in~\cite{Norr21}, we verify \mEdhoc{} in an extended
Dolev-Yao model~\cite{DY83}.
%
Dolev-Yao is a well-established model for symbolic verification of security
protocols.
%
Messages are modelled as terms in an algebra, and the various cryptographic
operations are assumed to be perfect, e.g., encrypted messages can only be
decrypted with the correct key, and there are no hash collisions exist.
%

The adversary is assumed to be the network, and thus controls the communication
channel and can see all messages being communicated as part of the protocol.
%
In addition, they can interact with an unbounded number of protocol sessions,
drop, inject and modify messages at will.
%

On top of the Dolev-Yao model,~\cite{Norr21} allow the adversary two extra
capabilities with respect to getting access to the long-term and ephemeral keys.
%
Long-term key reveal, denoted $\mRevLTK^{t}(A)$, stands for the adversary
gaining access to a party $A$'s long-term private key \mPriv{A} at a time point $t$.
%
Ephemeral key reveal, denoted $\mRevEph^{t}(A, k)$, stands for the adversary
obtaining, at time $t$, the ephemeral private key \mPrivE{A} used by party $A$
in a session where they establish key material $k$.
%
% Ephemeral keys might be stored in a less secure part of a device,
% but long-term keys are often stored and operate within a secure model,
% like a trusted execution environment (TEE). \vnote{Do we want to say more about TEEs here?}
%
Formalizing these two capabilities allows more fine-grained control
over the access that an adversary has to these fundamentally different kinds of
keys.
%

%---------------------------------------------------------------------- subsub
\subsubsection{System and Adversary Model Extensions Supporting TEE.}
\label{sec:TEE:advModel}
%
We extend the adversary model of~\cite{Norr21} by allowing the adversary
access to a TEE interface on a party of their choosing before and after the
\mEdhoc{} session.
%
This allows us to distinguish between a full compromise and a weak one, using
the terminology of~\cite{DBLP:conf/csfw/Cohn-GordonCG16}.
%
A weak compromise is when the adversary have temporary access to the TEE.
%
Seen in the framework of~\cite{DBLP:conf/icics/XuZRWTZ20}, we add adversary
capabilities corresponding to a server adversary, i.e., an adversary that
compromises a party, learns its ephemeral keys and may temporarily access its
TEE, but does not learn the party's long-term key.
%
The model of~\cite{Norr21} only captures server adversary capabilities without
access to the TEE, via the leakage of ephemeral keys, and local adversary
capabilities, or more specifically, a secure-world adversary with read-write
access to the
secure world where the long-term keys reside via the leakage of long-term keys.
%
Our extension allows analysis of so-called weak post-compromise security
(PCS)~\cite{DBLP:conf/csfw/Cohn-GordonCG16}.
%

We split the \mEdhoc{} functionality as follows.
%
The TEE contains the long-term key and allows the non-TEE parts of the
application to perform the operations on it via its interface.
%
More precisely, parties using the \mSig{} authentication method use a TEE with
an interface which accepts a message and returns the signature of that message
using the party's private long-term key.
%
A party $U$ using the \mStat{} authentication method use a TEE with an interface
accepting a point $P$ on the curve and returning $\mDH(P,\ \mGu{})$.
%

The non-TEE parts implement \mEdhoc{} using this interface.
%
This interface requires the least functionality from the TEE, reducing the TEE's
complexity and possibly cost of its implementation.
%
The functional split is suitable even for the situation where a
constrained device has implemented only the storage of the long-term key in a
special purpose circuit with minimal processing functionality.
%
Since \mEdhoc{} focuses on constrained IoT devices, it seems appropriate to
cater for this setting.
%

%---------------------------------------------------------------------- subsub
\subsubsection{Extended Formalism and Security Properties.}
\label{sec:TEE:fmAndProps}
Formally, we model the TEE interface by adding two new rewrite rules:
%
\begin{small}
\begin{verbatim}
rule forge_SIG:
   [!LTK_SIG($A, ~ltk), In(xx)] --[TEE($A)]-> [Out(sign(xx, ~ltk))]

rule exp_STAT:
   [!LTK_STAT($A, ~ltk), In('g'^~xx)] --[TEE($A)]-> [Out(('g'^~xx)^~ltk)]
\end{verbatim}
\end{small}
%
These rules allow the adversary to obtain terms representing signatures on a
value of their choice to forge signatures (\verb|forge_SIG|), or to obtain terms
representing a curve point of their choice raised to the power of the
long-term key (\verb|exp_STAT|).
%

Because it is a trivial attack when the adversary access these rules with values
from the test session, we must disqualify those rule applications.
%
We do so by creating an action fact \verb|TEE($A)|, where \verb|$A| is the
identity corresponding to the private key used, and then augmenting the
properties with a condition that no such action fact exists from the start of
the protocol execution and its end.
%
Care must be taken when specifying the start and the end.
%
Specifically, the start and end of the protocol run must be viewed w.r.t. the
role.
%
For example, the injective agreement property for the initiator in
the \mSigSig{} method requires that the adversary does not have access to the
TEE of the
responder from the time of transmission of the first message until the second
message is received by the initiator.
%
Because reception of one message and transmission of the next one at a party
is an atomic operation, these timepoints represent the start and end of the
protocol run from the perspective of the initiator.
%

We did not manage to get Tamarin to terminate for all lemmas and all methods
with the added TEE extension.
%
When the responder uses the \mStat{} authentication method Tamarin did not
complete the in a reasonable time for the authentication lemmas.
%
To combat this, we simplified the modeling of XOR encryption of the second
message for these lemmas when the responder uses \mStat{} authentication.
%
More precisely, the second message contains the two values \verb|R| (the
responder's identity) and
\verb|authR| (the authentication information), and we generally model the
XOR encryption by xoring each of them with their own key-stream terms.
%
For the problematic lemmas however, we resorted to xoring the entire tuple
\verb|<R, authR>| with a single key-stream term.
%
Given that no attacks were identified using our original modeling, this
simplification is intuitively not a severe restriction.
%
However, it is possible that we now miss attacks on these methods
combining properties of the two-term XOR encryption with TEE capabilities.
%

We now briefly describe \mTamarin{}, and how we model \mEdhoc{} in it here.
%

%-------------------------------------------------------------------------- sub
\subsection{\mTamarin{}}
\label{sec:tamarin}
We extend the formal \mEdhoc{} model of~\cite{Norr21}, using the symbolic
model tool \mTamarin{}~\cite{DBLP:conf/cav/MeierSCB13}, which is an interactive
tool for the formal verification of security protocols.
%
Protocols are modelled in \mTamarin{} as multiset rewrite rules which encode a
transition relation.
%
The elements of these multisets, called facts, contribute to the global
system state.
%
For syntactic sugar, \mTamarin{} also allows the use of let-bindings and tuples.
%
For ease of presentation, we will present the model and properties in a
slightly different syntax in this paper, but this syntax can be directly
mapped to that of \mTamarin.
%

Rewrite rules can be annotated with events, called actions in \mTamarin{}.
%
Communicated messages in the protocol are modelled as terms in an algebra,
which specifies sets of names, variables, and allowable function symbols.
%
Facts and actions are modelled as $n$-ary predicates in the term algebra,
and actions can be parametrized using terms.
%

An annotated multiset rewrite-rule is represented as
$l \ifarrow[e] r$, where $l$ and $r$ are multisets, and $e$ is
a multiset of actions.
%
A sequence of actions yields a protocol execution trace.
%
Event types are predicates over global states generated during protocol execution.
%
Consider an event type $E$ and a timestamp $t$ as part of a trace.
%
By $E^{t}(p_i)_{i\in\mathbb{N}}$, we represent an event of type $E$ occurring
at time $t$ in a trace, parametrized by the sequence of values
$(p_i)_{i\in\mathbb{N}}$
(corresponding to the action fact $E(p_i)_{i\in\mathbb{N}}@t$ in \mTamarin).
%
Thus, the time points form a quasi order, and we denote the fact that
$t_{1}$ comes before $t_{2}$ in a protocol trace by
$t_{1} \lessdot t_{2}$, and that $t_{1}$ and $t_{2}$ stand for the same
time point in a trace by $t_{1} \doteq t_{2}$.
%
\mTamarin{} allows events to occur at the same time point, with one
restriction: multiple events of the same type cannot occur simultaneously,
so if $t_{1} \doteq t_{2}$, then $E^{t_{1}} = E^{t_{2}}$.
%

Properties are defined as formulas in a fragment of temporal first order logic,
and these formulas can be verified over execution traces.
%

@@

Users can extend the default term algebra and equational theory in
\mTamarin{} with new function symbols and unification rules for these new symbols.
%
For example, \mEdhoc{} requires authenticated encryption, which we model using
the symbol \mT{aeadEncrypt}.
%
We augment \mTamarin{} with the following rule for this operation,
which represents the fact that if the adversary knows a key \mT{k},
a message \mT{m}, and authenticated data \mT{ad},
and has access to an encryption algorithm \mT{ai},
then they can obtain the message corresponding to the authenticated encryption
of \mT{m} with \mT{k}~\cite{Norr21}.
%
\begin{verbatim}
[!KU(k), !KU(m), !KU(ad), !KU(ai)] --[]-> [!KU(aeadEncrypt(k, m, ad, ai))]
\end{verbatim}
%
We make used of \mTamarin{}'s builtin rules for other operations, like \mXor{}.
%

The rules modeling a Dolev-Yao adversary and the evolution of their knowledge
as the protocol executes are built into \mTamarin.
%
We can extend and/or modify this behavior by adding other rules as well.
%
To denote that the adversary has access to a message $p$ at time $t$, we use $\mK^{t}(p)$.
%
This stands for the fact that as the adversary interacts with the parties
executing the protocol, at time point $t$, the adversary can derive the
term $p$ using the Dolev-Yao deduction rules and any additional adversary
capabilities specified in \mTamarin.
%
As an example, consider the following implication
\[
    \forall t, k, k'\mLogicDot \mK^{t}(\langle k, k'\rangle)\ \rightarrow\
\mK^{t}(k) \land \mK^{t}(k'),
\]
which represents the fact that if the adversary gets to know the pair of
keys $\langle k, k' \rangle$ at a time point $t$, then the adversary also
knows each of those keys $k$ and $k'$ at time point $t$ as well.
%
For more details about how \mTamarin{} manages adversary knowledge,
see~\cite{DBLP:conf/cav/MeierSCB13}.
%

%-------------------------------------------------------------------------- sub
\subsection{Model and Desired Properties}
\label{sec:desired-properties}
In this section, we describe how we model \mEdhoc{} and the various properties
we want it to satisfy.
%

@@

In this work, we consider the following properties:
perfect forward secrecy, injective agreement for both the initiator and the
responder, and implicit agreement.
%
Agreement is considered on a set of parameters $S$ which also contains the
session key material \mSessKey{}.
%
We will first describe in detail both properties, and then describe the
contents of the set $S$.
%
We formalize these properties as shown in Figure~\ref{fig:props}, which is
a modified version of a figure in~\cite{Norr21}.
\vnote{The figure changes because of TEEs}

@@

%----------------------------------------------------------------------- PFS
\subsubsection{Perfect Forward Secrecy (PFS).}
\label{sec:secrecy}
Informally, by perfect forward secrecy, we mean that the adversary cannot
gain access to the session key material \mSessKey{}, even if a long-term key is
leaked after the establishment of \mSessKey{}.
%
PFS for session key material \mSessKey{} is formalized as \mPredPfs{}
in Figure~\ref{fig:props}.
%

In the \mIComplete{} event, $I$ represents the identity of the initiator,
and $R$ represents the identity of the party who the initiator believes is
playing the responder role, while \mSessKey{} stands for the established
session key material.
%
\mRComplete{} has analogous parameters; in particular, the responder $R$
believes $I$ is the party playing the initiator role.
%
Intuitively, the PFS property states that an adversary obtains \mSessKey{}
only if a party's long-term key is compromised before their run ends, if an
ephemeral key is compromised at any time after a party's run began.
%
This is to hold even if the adversary has access to the TEE of either party
before and after the protocol execution.
%
The property in Figure~\ref{fig:props} is unlike the actual \mTamarin{} lemma
in one minor way: \mTamarin's logic does not allow disjunctions to appear on
the left-hand side of an implication inside a universally-quantified formula.
%
Therefore, in the \mTamarin{} code, instead of using the disjunction
$\mIComplete^{t_2}(I, R, \mSessKey)\, \lor\,  \mRComplete^{t_2}(I, R, \mSessKey)$
to model the fact that either party can have completed their execution, we use
a single action parametrized by the terms $I$, $R$, and \mSessKey.

%-------------------------------------------------------------------- InjAgree
\subsubsection{Authentication Properties.}
\label{sec:authenticationDef}
Following~\cite{Norr21}, we prove two different kinds of authentication
properties, namely \emph{injective agreement} in the style
of~\cite{DBLP:conf/csfw/Lowe97a}, and implicit agreement.
%
Injective agreement can be guaranteed to either party running the protocol.
%
For the initiator $I$, it guarantees to $I$ that whenever $I$ believes that
they have completed a run with $R$ as responder, then the party $R$ has indeed
executed the protocol in the role of a responder, and that this run of $I$
uniquely corresponds to one of $R$ where the set of parameters is $S$ and
includes, in particular, the session key material \mSessKey{}.
%
It can be defined for $R$ in a similar manner.

%
We formalize injective agreement for the initiator role as \mPredInjI{} and
for the responder role as \mPredInjR{} in Figure~\ref{fig:props}.
%
For the initiator $I$, this property represents the fact that either
injective agreement (as described above) holds, or the long-term key of
the party $R$ assumed to be playing the responder role has been
compromised before $I$'s role finished.
%
The property is to hold even if the adversary has access to the TEE of party $R$
before and after the protocol execution.
%
%If the adversary has managed to compromise $R$'s long-term key, they can
%generate any message of their liking and sign it using said key,
%or compute a $\mathit{MAC}_{R}$ using the long-term keys \mPubE{I},
%\mPriv{R}, and an ephemeral key pair $\langle\mPrivE{R},\ \mPubE{R}\rangle$
%of their choice.
%%
The analogous definition holds for the responder $R$.
%

@@

The ``implicit'' in the name of the property stands for the fact that a party
$A$ assumes that any party $B$ who has access to the session key material
\mSessKey{} must, in fact, be the intended party, and that if $B$ is honest,
$B$ will agree on a set $S$ of parameters which includes \mSessKey.
%
Implicit agreement for both roles guarantees to $A$ that $A$ is or has been
involved in exactly one protocol execution with $B$, and that $B$ agrees or
will agree with $A$ on $S$.
%
This property diverges from injective agreement in that upon sending
the last message, $A$ concludes that if this message reaches $B$, then $A$
and $B$ agree on each other's identities and roles, and the set $S$.
%

Note that for implicit agreement to hold for $I$, the ephemeral keys must not
be revealed since the property relies on the fact that the intended responder
is the only one who knows the session key material.
%
If the adversary has access to the ephemeral keys, they can use them along with
the public keys of $I$ and $R$ to compute the session key material.
%
However, either party's long-term key can be revealed after the other
party has finished their execution, since this still leaves the adversary
unable to compute \mGxy{}.
%

Since \mTamarin{} runs out of memory to verify this property as is,
we split it into two lemmas -- \mPredImpI{} for $I$ for one \mPredImpR{} for $R$.
%
Figure~\ref{fig:props} contains the definition for \mPredImpI{}.
%
\mPredImpR{} is formalized similarly, so we omit it.
%

%------------------------------------------------------- Agreed parameters
\subsubsection{Parameters in set $S$.}
\label{sec:agreedParams}
We now describe the set $S$ of parameters upon which the two parties obtain
guarantees via the above properties.
%
The initiator $I$ gets injective and implicit agreement guarantees on the
following partial set $S_P$ of parameters~\cite{Norr21}:
\begin{itemize}
    \item the roles played by itself and its peer,
    \item responder identity,
    \item session key material (which varies depending on the \mEdhoc{} method),
    \item context identifiers \mCi{} and \mCr{}, and
    \item cipher suites \mSuites{}.
\end{itemize}
%

@@

In addition to these properties, a couple of properties can be inferred
without being explicitly modelled and verified.
%
One such property is Key-Compromise Impersonation
(KCI)~\cite{DBLP:conf/ima/Blake-WilsonJM97}.
%
A KCI attack occurs when an adversary who has access to $A$'s long-term private
key to make $A$ believe that they completed an execution with a peer $B$,
while $B$ did not participate in said execution at all.
%
This is in particular relevant when \mStat{} authentication methods are used.
%
Our above notions of agreement ensure that both parties agree on each
other's identity, role, and session key material.
%
Therefore, all \mEdhoc{} methods that satisfy these agreement properties also
avoid KCI attacks.
%

Another kind of attack is Unknown Key Share attacks
(UKS)~\cite{DBLP:conf/ima/Blake-WilsonJM97}.
%
As part of a UKS attack, a party $A$ can be made to believe that it finished
an execution with a party $B$, but where the session key material is actually
shared between $A$ and $C$ instead.
%
Again, due to the agreement on identities and session key material, any method
that satisfies the above agreement properties also resists UKS attacks.
%
Overall, the agreement properties capture entity authentication,
and satisfy any properties based on that notion.
%
However, see Section~\ref{sec:unintendedPeerAuth} for a discussion on the
interaction between \mEdhoc{} and the application leading to similar issues.
%

%---------------------------------------------------------------------------sub
\subsection{Encoding \mEdhoc{} in \mTamarin}
\label{sec:modeling}
%
In this section, we describe how we model the \mSigSig, \mSigStat, \mStatSig,
and \mStatStat{} methods of \mEdhoc{} in \mTamarin.
%
As in~\cite{Norr21}, we construct the \mTamarin{} model by utilizing the fact
that all methods of \mEdhoc{} share a common underlying structure
(as shown in Figure~\ref{fig:edhocFramework}).
%
We do so by using a single specification file in the M4 macro language,
which generates all the methods.
%
As in~\cite{Norr21}, we only present the \mStatSig{} method which illustrates
the use of two different authentication methods.
%
The full \mTamarin{} code for all models can be found at~\cite{edhocTamarinRepo}.
%

@@

In addition to these default operations, \mEdhoc{} is built over
\mHkdfExpand, \mHkdfExtract, and the \mAead{} functions.
%
We represent \mHkdfExpand{} by \mT{expa} and \mHkdfExtract{} by \mT{extr}.
%
\mAead{} operations need us to add extra equations to the underlying theory.
%
A term encrypted using \mAead{} is represented by \mT{aeadEncrypt(m, k, ad, ai)},
where \mT{m} is the underlying message, \mT{k} is the encrypting key,
\mT{ad} is the additional data, and \mT{ai} is the identifier for the
encryption algorithm.
%
Decryption of such a term is defined via two equations that we add to
\mTamarin's theory.
%
The following equation requires the decrypting party to know the additional
data \mT{ad} to decrypt this encrypted term with verification of its integrity.
\begin{small}
\begin{verbatim}
aeadDecrypt(aeadEncrypt(m, k, ad, ai), k, ad, ai) = m.
\end{verbatim}
\end{small}
%
Only the above equation is used by honest parties, but the adversary should
also be able to decrypt without having to go through the additional step of
identity verification.
%
To this end, we also add the following equation, where the adversary does not
need access to the additional data \mT{ad} in order to decrypt.
%
\begin{small}\begin{verbatim}
decrypt(aeadEncrypt(m, k, ad, ai), k, ai) = m.
\end{verbatim}\end{small}
%

@@

To model the reveal of long-term keys and ephemeral keys to an adversary,
we use standard reveal rules and events of type \mRevLTK{} and \mRevEph,
respectively.
%
It is also important to keep track of the time points at which these events occur.
%
Long-term keys can be revealed on registration, even before the protocol begins.
%
Ephemeral keys in our model can only be revealed when a party
completes their role, i.e., simultaneously with events of type \mIComplete{}
and \mRComplete.
%
Having set out the capabilities of the adversary, we now model the execution
of the honest agents' roles.
%

For each protocol method, we use two rules apiece for the initiator and
responder -- \mT{I1}, \mT{R2}, \mT{I3}, \mT{R4}.
%
Each of these stand for one step of the protocol as executed by either party.
%
To disambiguate, we will attach the method to the rule name.
%
These four rules directly map to the event types
\mIStart, \mRStart, \mIComplete, and \mRComplete, respectively.
%
We show the \mT{R2\_STAT\_SIG} rule below to illustrate the various aspects
of the \mTamarin{} modeling we are describing here.
%
